Combination of field and experimental data together with computational models reveal cognitive mechanisms behind cleaner fish behaviour

### 2022-02-22 manuscript_1.0/manuscript_1.0.tex
Error found in the manual coding of the cleaner entries used to compute visitor preference. In loadFieldData.R a programatic way of coding cleaner entries was done. Now that the manual and programatic results coincide I can re-run the chains. New figure configurations were implemented. Among those supplementary figures showing the diagnostics of the MCMC. A new result section was written.

%%%% Abstract text to be placed here %%%%%%%%%%%%
\begin{abstract}
While it is generally straightforward to quantify individual performance in cognitive experiments, identifying the underlying cognitive processes remains a major challenge. Often, different mechanistic underpinnings yield similar performances, and Lloyd Morgan's cannon warrants acceptance of the simpler explanation. Alternatively, when the different mechanisms interact with environmental conditions, variation in performance across environments might allow to statistically infer the mechanism responsible. However, to do this it is necessary to have quantitative predictions for the different candidate mechanisms. Here, we use a set of computational models to get quantitative predictions from alternative learning mechanisms, and using Bayesian statistics fit the model parameters to performance data. We used experimental data on performance in an ephemeral reward task by wild-caught cleaner fish \emph{Labroides dimidiatus}, as well as cleaner and client fish densities from the locations of capture. The task can in principle be solved by estimating future consequences of an action, or by perceiving the removal of the ephemeral reward as psychological punishment (negative reinforcement). We found that a model where cleaners only estimate the future consequences of their actions explains best which cleaner relative abundances cause the fish to develop a preference for an ephemeral food source. This model also yields performances that can be considered the result of locally optimal decision-rules, in contrast to the negative reinforcement model. We argue that the combination of computational models with data is a powerful tool to infer the mechanistic underpinning behind cognitive performance.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

When researchers aimed at testing wild-caught cleaner fish, as
well as individuals from other species, abilities
to prioritize visitor over resident clients in a lab-based paradigm,
they used Plexiglas plates of different colours and shapes offering same
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e.~it remains until
the cleaner had eaten off it. The other plate was made to behave
like a visitor, i.e.~it is removed if not inspected first.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food \citep{bshary_Asymmetric_2002}.
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the biological market task
\citep{salwiczek_Adult_2012, zentall_Early_2017}; yet the African grey
parrots solve this task as well \citep{pepperberg_Can_2014}. Given that primates
are known to readily distinguish
between one and two reward units if presented in more conventional ways
\(\color{red}{\text{ref}}\), this raises the question:
what makes the task difficult for most non-cleaners?

@@

The cognitive performance data was from a total of 120 cleaners
(10 individuals per 12 site/year) tested in the
ephemeral reward task \citep{triki_Biological_2019, triki_Brain_2020}. Authors
housed all captured cleaners individually in glass aquaria
( \(62cm \times 27cm \times 37 cm\) ) and provided them
with PVC pipes (\(10 cm \times 1 cm\)) as shelters.
The task consisted of exposing the cleaner fish to substitute
models of client fish in the form of two \emph{Plexiglas} plates offering the
same amount of food (one item of mashed prawn). The two plates differed
in colour and pattern (horizontal green stripes or vertical pink stripes)
but had equal size (\(10 cm \times 7 cm\)). Importantly, the two plates played
different roles as either a visitor (ephemeral food source) or
resident (permanent food source). That is, if a cleaner inspected the
resident plate first, the experimenter withdrew the visitor plate out of
the aquarium as a consequence. Choosing first the visitor plate,
however, granted access to both plates. The equal size of the plates
forced cleaner fish to learn to give service priority to the visitor plate
based solely on the behaviour-cue of the plates rather than size-cue
\citep{wismer_Cuebased_2019}. Triki \emph{et al.} \citep{triki_Biological_2019, triki_Brain_2020}
tested the fish for a maximum of 200 trials with 20 trials a day, 10 trials
in the morning and 10 trials in the afternoon. They randomized and
counterbalanced the plates' spatial location (i.e.~left or right)
between trials. Similarly, they counterbalanced the plates' decoration
(colour and pattern) and the plates' role (visitor or resident) between the
tested fish. In the original studies, once a fish reach a learning criterion,
that is, performing significantly above chance level (\(> 50%
\), \(p-value ≥ 0.05\)),
they passed to a reversal version of the task where the roles of the
visitor/resident Plexiglas plates were swapped (see Triki et al.~2019, 2020).
Here, we used instead a subset of these data in order to have an idea on
cleaner fish preferences for the visitor plate, even if they do not reach
the learning criterion within 200 trials. To do so, we first extracted the
trial-by-trial outcomes from the last two session (20 trials) of those who
never reached the learning criterion for visitor plate
\(\color{red}{\text{(N = XX cleaner fish)}}\).
For those who reached the learning criterion at some point during the
test and passed to a reversal phase, we extracted the trials-by-trial
outcomes from the last session (10 trials) before passing to reversal
and the last 10 trials they were exposed to in the test
\(\color{red}{\text{(N = XX cleaner fish)}}\).

@@

The aim of the analyses is to fit the key model parameters
\(\gamma\) and \(\eta\), to the empirical data from Triki \emph{et al.}
\citep{triki_Biological_2019, triki_Brain_2020} to test whether each or
a combination of these effects is a better explanation for the pattern
seen in the data. We used the
ecological variables: cleaners, visitor clients,
resident clients abundances and
visitor clients leaving probability, as input to the models.
As response variable we used cleaners' preference for visitor
over residents in the ephemeral reward task.
Finally, we used the preference for visitors resulting from the
model simulations as the prediction for the response variable.
We kept all other parameter values used for the model simulations constant,
see Table \ref{tab:param}.

To capture with the model the relationship between the ecological
variables and cleaner fish preferences for visitors we need to scale
the absolute population densities of cleaner fish from the empirical data
to a measure of relative abundances that captures
client visitation patterns. This is because in the model, relative
abundances of clients define not only the probability of residents and visitors,
but also how often the cleaning station is empty (e.g.~there are no
clients to be cleaned). The frequency with which clients visit the station
is another variable, which in nature may vary among different client species
depending on their ectoparasite loads. We do not have field estimates
for species-specific parasite loads, especially not as a function of site.
In order to control for these aspects, we computed a measure of relative
cleaner fish abundance for each reef site relative to
absolute abundances and multiplied it
by a scaling constant that changes the range of the variable. This scaling
constant is hence meant to capture variation in the market conditions driven
partly by cleaner abundance. We fitted the value of the scaling constant as
part of the statistical inference. As for the visitor and resident
abundances we computed a relative measure with respect to the
total client abundance, and weighted that by the rescaled relative absence
of cleaners (\(1-\text{relative cleaner abundance}\)). Thus, all three measures
of relative abundance sum up to 1, and can be used in the model as a proxy
for the probability of having different options in the cleaning station.
Note that we introduced the scaling constant to control for variation that
is not captured by the model; its parameter distribution does not offer
biological insights.

Once we calculated the relative abundances, we obtained predictions from
the model for each one of the locations and ran the Markov Chain. We started the
chain with random values for the parameters of interest, then ran the
computational model once for every reef site using as input the ecological
explanatory variables. The model outputs the probability of choosing the visitor
for each location \(p_i\), where \(i\) is the index for the 12 locations. Assuming a
binomial distribution, the probability that each of the cleaner fish 20 choices
in the ephemeral reward task was generated by the model
is given by \(\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}\);
where \(k_{ij}\) is the number of times that cleaner fish \(j\) in reef site \(i\)
chose the visitor over the resident; and \(n=20\) (due to our choice of using 20
choices per cleaner fish). By taking the natural logarithm and summing over all
the cleaner fish and reef sites we obtained the log-likelihood of the data
given the model and parameters. We then proposed a new set of parameters drawn
from a uniform distribution centred around the old parameter set. The amplitude
of the uniform distribution used for each parameter can be found in Table
\ref{tab:param}. Subsequently, we ran the model and calculated the likelihood
with the new parameter set. We then used the ratio of the two likelihoods to
choose which parameter set to keep. New parameter sets with a higher likelihood
than the old set replaced old ones, and those with a lower likelihood replaced
current ones with a probability equals to the log-likelihood ratio.
Given that we only used the likelihoods in the decision, we used a
uninformative prior. Once we decided whether the new parameter set replaces
the old one, we ran the model again to sample the likelihood
distribution of the parameter set. We then started
the cycle again by proposing a new set of parameters, and repeated
the process for \(1e^5\) steps. We ran 5 independent chains,
discarded the first \(100\) samples of each chain as burn in,
and after that we kept 1 in every 100 samples to avoid autocorrelation.
The collection of parameters sets kept in
all chains approximates the posterior distribution. We coded the model
as well as the fitting algorithm in \emph{c++}, the diagnostics and visualization
in R \citep{rcoreteam_Language_2021}. All codes are accessible at
\(\color{red}{\text{include citation}}\)

@@

{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/post-1}

\caption{Posterior distributions for parameter values $\gamma$, $\eta$ and the scaling constant. We show the kernel density estimates, below the mode (black dot) and the 65\% (light blue shade) and 95\% (grey shade)  highest posterior density interval for the three paramteres. On the left, panels a, b and c show posterior distributions for a model with all three parameters; while on the right, d and e, show a model with only chaining. Panel f shows a measure of fit for both models, namely the distribution of pseudo-$R^2$ obtained from sampling the posterior distribution of parameter values.}\label{fig:post}
\end{figure}

@@

{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/pred-1}

\caption{Observed and predicted probability of choosing a visitor. Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 reef sites, as well as the corresponding relative cleaner fish abundance (x axis) and frequency of visitors clients leaving the cleaning station (y axis). Right-hand side panels: Variation of the predicted probabilites  of choosing a visitor over a resident and their observed values for 12 locations. Circles show the mean prediction for each location from 100 samples taken from the posterior distribution. Thick and thin bars show the 66 and 95\% credible interval, respectively, taken from those posterior samples. Squares show the predictions used for panel on the righ-hand side. Colour coding denotes different reef sites/year of the data collection (see Empirical data section). Black line corresponds to a perfect match between observed and predicted probabilities. Upper panels show prediction from a model including chaining and operant conditioning, lower panels from a  model with only chaining.}\label{fig:pred}
\end{figure}

@@

While all three models captured well the positive relation between visitor
leaving behaviour and cleaner performance in
the market task \citep{triki_Biological_2019}, only the chaining mechanism
predicted that cleaner fish performance in the task should be low in
habitats with low cleaner-to-client ratios, regardless of the
visitor leaving probability. In contrast, models including
negative reward predicted the highest performance in the ephemeral
reward task when relative cleaner abundance is low, particularly
together with high probability of visitor leaving (Fig. \ref{fig:pred}).
Low relative cleaner fish abundances mean the market has an excess of demand for
cleaning services. In the models, this translates to a cleaning station
that is frequently full. Thus, when a visitors leaves, it is likely that the
cleaner fish will have access to another client in the next step. Therefore,
there will not be much difference in future reward between choosing
a visitor and a resident, and cleaners will not develop a preference for the
visitor in these conditions. On the other hand, the effect of negative reward
on cleaner fish preference is the opposite, as in a busy cleaning station, to
that of chaining. cleaner fish will get more often the resident-visitor
state and will develop a preference for the visitor faster.
At high cleaner fish abundances, the resident-visitor state
becomes so rare that neither mechanism is very efficient
at generating a preference for visitors. When facing the resident
visitor choice it is still be best to choose the visitor; however,
the learning machinery will not be able to develop this preference
efficiently. Overall, the models suggest that chaining is the cognitive
mechanism that allows cleaner fish to adaptively adjust to their
biological market ecological conditions.

Our results support chaining as the more plausible mechanism behind the
inter-specific variation in cleaners' performance. This suggest
that chaining could be under positive selection. In the standard ephemeral
reward task, however, food amount and quality are equal in the
two Plexiglas plates. Since our computational models simulate this aspect of
the task, choosing visitors over residents at low cleaner fish abundances
does not diminish the amount of food/reward in the models.
On the other hand, a cleaner fish committed to visitors might find it more difficult
to include in its decision making other variables, like parasite load
and/or mucus quality \citep{roche_Client_2021}. Indeed, in the absence of
market effects, both field observations and aboratory experiments also
using simultaneous two choice options, show that cleaner fish base their
decision-making on other cues such as the mucus quality and the amount of food.\\
\citep{triki_Fluctuations_2019, grutter_Cleaner_2004}, as well as client size
as a proxy for food quantity \citep{wismer_Cuebased_2019, grutter_Does_2005}.
How important this limitation is will determine whether chaining yields a higher
fitness for cleaners.

Previous research showed that cleaner fish living at high population
densities and giving service priority to the visitor plate in the
ephemeral reward task, as well as cleaner fish living at low
densities but denying service priority to the visitor plate possess
larger forebrains; a key teleost brain region associated with
behavioral flexibility and social intelligence. Those failing to
show optimized decision-rules given their local ecological
conditions had relatively smaller forebrains \citep{triki_Brain_2020}.
Triki \emph{et al.} refer to the former as socially competent cleaner fish,
while the second group as socially incompetent cleaner fish.
Social competence is the ability to optimise social behaviour
to the available social information \citep{taborsky_Social_2012a, bshary_Cooperation_2015, varela_Correlated_2020}. Our analyses yielded no
evidence that the difference in social competence with respect to the local
ecological conditions and associated brain morphology,
found by Triki \emph{et al} \citep{triki_Brain_2020}, are due to the mechanism
used to incorporate future consequences. It could have been conceivable that
high performing individuals from low population densities reef sites
might use negative reinforcement instead of chaining, but in that
case negative reinforcement should have explained at least
part of the data. Configurational learning or
chunking \citep{sutherland_Configural_1989, miller_Magical_1956},
the second component necessary to solve the ephemeral
reward task \citep{quinones_Reinforcement_2019}, was
not varied in the models we analized here. However, while chunking tendencies
should vary to allow individuals to adapt to their local conditions
\citep{prat_Modelling_2022, kolodny_Evolution_2014}, systematic differences
in individual chunking tendencies would not explain how socially
competent decisions vary as a function of relative abundance.
Therefore, it remains currently unclear what cell-demanding
mechanisms may cause variation in social competence that
translates into site-specific variation in performance in the
ephemeral reward task.

@@

@@

{\centering \includegraphics[width=0.95\linewidth,]{manuscript_1.0_files/figure-latex/nogamma-1}

@@

\caption{Observed and predicted probability of choosing a visitor for a model without chaining. Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 reef sites, as well as the corresponding relative cleaner fish abundance (x axis) and frequency of visitors leaving the station if being made to wait (y axis). Right-hand side panel: Right-hand side panels: Variation of the predicted probabilites  of choosing a visitor and their observed values for 12 reef sited. Circles show the mean prediction for each location from 100 samples taking from the posterior distribution. Thick and thin bars show the 66 and 95\% credible interval taken from those posterior samples. Squares show the predictions used for panel on the righ-hand side. Colour denotes different locations (see Empirical data section). Black line correspond to a perfect match between observed and predicted probabilities.}\label{fig:nogamma}
\end{figure}

@@

>>>>

%%%% Abstract text to be placed here %%%%%%%%%%%%
\begin{abstract}
While it is generally straightforward to quantify individual performance in cognitive experiments, identifying the underlying cognitive processes remains a major challenge. Often, different mechanistic underpinnings yield similar performances, and Lloyd Morgan's cannon warrants acceptance of the simpler explanation. Alternatively, when the different mechanisms interact with environmental conditions, variation in performance across environments might allow to statistically infer the mechanism responsible. However, to do this it is necessary to have quantitative predictions for the different candidate mechanisms. Here, we use a set of computational models to get quantitative predictions from alternative learning mechanisms, and using Bayesian statistics fit the model parameters to performance data. We used experimental data on performance in an ephemeral reward task by wild-caught cleaner fish \emph{Labroides dimidiatus}, as well as cleaner and client fish densities from the locations of capture. The task can in principle be solved by estimating future consequences of an action, or by perceiving the removal of the ephemeral reward as psychological punishment (negative reinforcement). We found that a model where cleaners only estimate the future consequences of their actions explains best which cleaner fish relative abundances cause the fish to develop a preference for an ephemeral food source. This model also yields performances that can be considered the result of locally optimal decision-rules, in contrast to the negative reinforcement model. We argue that the combination of computational models with data is a powerful tool to infer the mechanistic underpinning behind cognitive performance.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

When researchers aimed at testing wild-caught cleaner fish, as
well as individuals from other species, abilities
to prioritize visitor over resident clients in a lab-based paradigm,
they used Plexiglas plates of different colours and shapes offering same
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e.~it remains until
the cleaner fish had eaten off it. The other plate was made to behave
like a visitor, i.e.~it is removed if not inspected first.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food \citep{bshary_Asymmetric_2002}.
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the biological market task
\citep{salwiczek_Adult_2012, zentall_Early_2017}; yet the African grey
parrots solve this task as well \citep{pepperberg_Can_2014}. Given that primates
are known to readily distinguish
between one and two reward units if presented in more conventional ways
\(\color{red}{\text{ref}}\), this raises the question:
what makes the task difficult for most non-cleaners?

@@

The cognitive performance data was from a total of 120 cleaners
(10 individuals per 12 site/year) tested in the
ephemeral reward task \citep{triki_Biological_2019, triki_Brain_2020}. Authors
housed all captured cleaners individually in glass aquaria
( \(62cm \times 27cm \times 37 cm\) ) and provided them
with PVC pipes (\(10 cm \times 1 cm\)) as shelters.
The task consisted of exposing the cleaner fish to substitute
models of client fish in the form of two \emph{Plexiglas} plates offering the
same amount of food (one item of mashed prawn). The two plates differed
in colour and pattern (horizontal green stripes or vertical pink stripes)
but had equal size (\(10 cm \times 7 cm\)). Importantly, the two plates played
different roles as either a visitor (ephemeral food source) or
resident (permanent food source). That is, if a cleaner fish inspected the
resident plate first, the experimenter withdrew the visitor plate out of
the aquarium as a consequence. Choosing first the visitor plate,
however, granted access to both plates. The equal size of the plates
forced cleaner fish to learn to give service priority to the visitor plate
based solely on the behaviour-cue of the plates rather than size-cue
\citep{wismer_Cuebased_2019}. Triki \emph{et al.} \citep{triki_Biological_2019, triki_Brain_2020}
tested the fish for a maximum of 200 trials with 20 trials a day, 10 trials
in the morning and 10 trials in the afternoon. They randomized and
counterbalanced the plates' spatial location (i.e.~left or right)
between trials. Similarly, they counterbalanced the plates' decoration
(colour and pattern) and the plates' role (visitor or resident) between the
tested fish. In the original studies, once a fish reach a learning criterion,
that is, performing significantly above chance level (\(> 50%
\), \(p-value ≥ 0.05\)),
they passed to a reversal version of the task where the roles of the
visitor/resident Plexiglas plates were swapped (see Triki et al.~2019, 2020).
Here, we used instead a subset of these data in order to have an idea on
cleaner fish preferences for the visitor plate, even if they do not reach
the learning criterion within 200 trials. To do so, we first extracted the
trial-by-trial outcomes from the last two session (20 trials) of those who
never reached the learning criterion for visitor plate
(\(N = 45\) cleaner fish).
For those who reached the learning criterion at some point during the
test and passed to a reversal phase, we extracted the trials-by-trial
outcomes from the last session (10 trials) before passing to reversal
and the last 10 trials they were exposed to in the test
(\(N = 75\) cleaner fish).

@@

The aim of the analyses is to fit the key model parameters
\(\gamma\) and \(\eta\), to the empirical data from Triki \emph{et al.}
\citep{triki_Biological_2019, triki_Brain_2020} to test whether each or
a combination of these effects is a better explanation for the pattern
seen in the data. We used the ecological variables: cleaners, visitor clients,
resident clients abundances and visitor clients leaving probability,\\
as input to the models. As response variable we used cleaners'
preference for visitor over residents in the ephemeral reward task.
Finally, we used the preference for visitors resulting from the
model simulations as the prediction for the response variable.
We kept all other parameter values used for the model simulations constant,
see Table \ref{tab:param}.

To capture with the model the relationship between the ecological
variables and cleaner fish preferences for visitors we need to scale
the absolute population densities of cleaner fish from the empirical data
to a measure of relative abundances that captures client visitation patterns.
This is because in the model, relative abundances of clients
define not only the probability of residents and visitors,
but also how often the cleaning station is empty (e.g.~there are no
clients to be cleaned). The frequency with which clients visit the station
is another variable, which in nature may vary among different client species
depending on their ectoparasite loads. We do not have field estimates
for species-specific parasite loads, especially not as a function of site.
In order to control for these aspects, we computed a measure of relative
cleaner fish abundance for each reef site relative to
absolute abundances and multiplied it
by a scaling constant that changes the range of the variable. This scaling
constant is hence meant to capture variation in the market conditions driven
partly by cleaner fish abundance. We fitted the value of the scaling constant as
part of the statistical inference. As for the visitor and resident
abundances we computed a relative measure with respect to the
total client abundance, and weighted that by the rescaled relative absence
of cleaners (\(1-\text{relative cleaner abundance}\)). Thus, all three measures
of relative abundance sum up to 1, and can be used in the model as a proxy
for the probability of having different options in the cleaning station.
Note that we introduced the scaling constant to control for variation that
is not captured by the model; its parameter distribution does not offer
biological insights.

Once we calculated the relative abundances, we obtained predictions from
the model for each one of the locations and ran the Markov Chain. We started the
chain with random values for the parameters of interest, then ran the
computational model once for every reef site using as input the ecological
explanatory variables. The model outputs the probability of choosing the visitor
for each location \(p_i\), where \(i\) is the index for the 12 locations. Assuming a
binomial distribution, the probability that each of the cleaner fish 20 choices
in the ephemeral reward task was generated by the model
is given by \(\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}\);
where \(k_{ij}\) is the number of times that cleaner fish \(j\) in reef site \(i\)
chose the visitor over the resident; and \(n=20\) (due to our choice of using 20
choices per cleaner fish). By taking the natural logarithm and summing over all
the cleaner fish and reef sites we obtained the log-likelihood of the data
given the model and parameters. We then proposed a new set of parameters drawn
from a uniform distribution centred around the old parameter set. The amplitude
of the uniform distribution used for each parameter can be found in Table
\ref{tab:param}. Subsequently, we ran the model and calculated the likelihood
with the new parameter set. We then used the ratio of the two likelihoods to
choose which parameter set to keep. New parameter sets with a higher likelihood
than the old set replaced old ones, and those with a lower likelihood replaced
current ones with a probability equals to the log-likelihood ratio.
Given that we only used the likelihoods in the decision, we used a
uninformative prior. Once we decided whether the new parameter set replaces
the old one, we ran the model again to sample the likelihood
distribution of the parameter set. We then started
the cycle again by proposing a new set of parameters, and repeated
the process for \(1e^5\) steps. We ran 5 independent chains,
discarded the first \(1000\) samples of each chain as burn in,
and after that we kept 1 in every 100 samples to avoid autocorrelation.
The collection of parameters sets kept in
all chains approximates the posterior distribution. We coded the model
as well as the fitting algorithm in \emph{c++}, the diagnostics and visualization
in R \citep{rcoreteam_Language_2021}. All codes are accessible at
\(\color{red}{\text{include citation}}\)

@@

@@

@@

While all three models captured well the positive relation between visitor
leaving behaviour and cleaner fish performance in
the market task \citep{triki_Biological_2019}, only the chaining mechanism
predicted that cleaner fish performance in the task should be low in
habitats with low cleaner-to-client ratios, regardless of the
visitor leaving probability. In contrast, models including
negative reward predicted the highest performance in the ephemeral
reward task when relative cleaner fish abundance is low, particularly
together with high probability of visitor leaving (Fig. \ref{fig:pred}).
Low relative cleaner fish abundances mean the market has an excess of demand for
cleaning services. In the models, this translates to a cleaning station
that is frequently full. Thus, when a visitors leaves, it is likely that the
cleaner fish will have access to another client in the next step. Therefore,
there will not be much difference in future reward between choosing
a visitor and a resident, and cleaners will not develop a preference for the
visitor in these conditions. On the other hand, the effect of negative reward
on cleaner fish preference is the opposite, as in a busy cleaning station, to
that of chaining. cleaner fish will get more often the resident-visitor
state and will develop a preference for the visitor faster.
At high cleaner fish abundances, the resident-visitor state
becomes so rare that neither mechanism is very efficient
at generating a preference for visitors. When facing the resident
visitor choice it is still be best to choose the visitor; however,
the learning machinery will not be able to develop this preference
efficiently. Overall, the models suggest that chaining is the cognitive
mechanism that allows cleaner fish to adaptively adjust to their
biological market ecological conditions.

Our results support chaining as the more plausible mechanism behind the
intra-specific variation in cleaners' performance. This suggest
that chaining could be under positive selection. In the standard ephemeral
reward task, however, food amount and quality are equal in the
two Plexiglas plates. Since our computational models simulate this aspect of
the task, choosing visitors over residents at low cleaner fish abundances
does not diminish the amount of food/reward in the models.
On the other hand, a cleaner fish committed to visitors might find it more difficult
to include in its decision making other variables, like parasite load
and/or mucus quality \citep{roche_Client_2021}. Indeed, in the absence of
market effects, both field observations and laboratory experiments also
using simultaneous two choice options, show that cleaner fish base their
decision-making on other cues such as the mucus quality and the amount of food.\\
\citep{triki_Fluctuations_2019, grutter_Cleaner_2004}, as well as client size
as a proxy for food quantity \citep{wismer_Cuebased_2019, grutter_Does_2005}.
How important this limitation is will determine whether chaining yields a higher
fitness for cleaners.

Previous research showed that cleaner fish living at high population
densities and giving service priority to the visitor plate in the
ephemeral reward task, as well as cleaner fish living at low
densities but denying service priority to the visitor plate possess
larger forebrains; a key teleost brain region associated with
behavioural flexibility and social intelligence. Those failing to
show optimized decision-rules given their local ecological
conditions had relatively smaller forebrains \citep{triki_Brain_2020}.
Triki \emph{et al.} refer to the former as socially competent cleaner fish,
while the second group as socially incompetent cleaner fish.
Social competence is the ability to optimise social behaviour
to the available social information \citep{taborsky_Social_2012a, bshary_Cooperation_2015, varela_Correlated_2020}. Our analyses yielded no
evidence that the difference in social competence with respect to the local
ecological conditions and associated brain morphology,
found by Triki \emph{et al} \citep{triki_Brain_2020}, are due to the mechanism
used to incorporate future consequences. It could have been conceivable that
high performing individuals from low population densities reef sites
might use negative reinforcement instead of chaining, but in that
case negative reinforcement should have explained at least
part of the data. Configurational learning or
chunking \citep{sutherland_Configural_1989, miller_Magical_1956},
the second component necessary to solve the ephemeral
reward task \citep{quinones_Reinforcement_2019}, was
not varied in the models we analized here. However, while chunking tendencies
should vary to allow individuals to adapt to their local conditions
\citep{prat_Modelling_2022, kolodny_Evolution_2014}, systematic differences
in individual chunking tendencies would not explain how socially
competent decisions vary as a function of relative abundance.
Therefore, it remains currently unclear what cell-demanding
mechanisms may cause variation in social competence that
translates into site-specific variation in performance in the
ephemeral reward task.

@@

{\centering \includegraphics{manuscript_1.0_files/figure-latex/rawdata-1}

\caption{Relation between the response variable used in this study and the
criteria used in previous studies to assess performance in the ephemeral reward
task. In the x axis, we classified the performance of cleaner fish according to
whether they developed a preference for the visitor in the initial round,
in the initial and reversal, or none of them. In the y axis, we add the choices
of two experimental sessions: panel on the left uses one session from the initial
round and one from the reversal round when possible
(as described in the main text); panel on the right uses two sessions from the
initial round for all fish.}\label{fig:rawdata}
\end{figure}

@@

{\centering \includegraphics[width=0.95\linewidth,]{manuscript_1.0_files/figure-latex/scaConst-1}

@@

\caption{Posterior distributions for scaling constant for the three model
(``Full model'', chaining only'' and ``penalty only''). We show the kernel
density estimates, below the mode (black dot) and the 65\% (light blue shade)
and 95\% (grey shade) highest posterior density interval. On the top,
panel a shows the posterior distribution from the full model; panel b from
the model with only chaining; and panel c from a model with only penalty.}\label{fig:scaConst}
\end{figure}

@@

\begin{verbatim}
## pdf 
##   2
\end{verbatim}

{\centering \includegraphics{manuscript_1.0_files/figure-latex/diagnosticsfull-1}

\caption{MCMC convergence diagnostics for the full model.
On the left trace-plots, on the right changes along the chain of the
Gelman and Rubin's shrink factor \citep{brooks_General_1998}.}\label{fig:diagnosticsfull}
\end{figure}

\begin{verbatim}
## pdf 
##   2
\end{verbatim}

{\centering \includegraphics{manuscript_1.0_files/figure-latex/diaggam-1}

\caption{MCMC convergence diagnostics for the chaining model.
On the left trace-plots, on the right changes along the chain of the
Gelman and Rubin's shrink factor \citep{brooks_General_1998}}\label{fig:diaggam}
\end{figure}
